---
title: "Debriefing_#3_March_31_Riu"
output: 
  html_document:
    theme: yeti
    toc: TRUE
    toc_depth: 5
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r Loading Libraries}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(lubridate)
library(gridExtra)
library(grid)
library(formattable)
library(zoo)
library(tibble)
library(forcats)
```


```{r Source}
source("dogs_cleaning.R")
source("cattle_cleaning.R")
source("sq_monkey_cleaning.R")
source("dogs_columns_addition.R")
source("cattle_columns_addition.R")
source("sq_monkey_columns_addition.R")
```

\

### **Goals**

1. **Work with the Cattle data and answer the client's questions regarding the animal**
    + Measured by the creation of visualizations to send to the client
    
2. **Work with Squirrel Monkey data and answer the client’s questions regarding the animal**
    + Measured by the creation of visualizations to send to the client
    
3. **Explore the differences in the structure of the animal datasets that we currently have.**
    + ~~Measured by the ability to articulate the differences during the next debriefing~~
    + Measured by the creation of a table accompanied by bullet point comments, both detailing the various differences 
    
4. **Create a prototype R Shiny program for only the Painted African Dogs, Cattles, and Squirrel Monkeys** 
    + ~~Measured by a live demonstration during the next debriefing~~
    + Measured by the posting of a link for people to test out the prototype R Shiny Program


* We worked on the four goals listed above for Debriefing #3
* For Goal 3 and Goal 4, we made slight modifications on how our progress can be measured, as debriefings are now deilivered in a written format


\

#### **Stereotypic Behaviors**
*How often are they showing Stereoptypic behaviors?*

```{r Table for Stereotypic Behaviors}

cattle_stereotypic <- cattle_data %>% filter(Category == "Stereotypic") %>%
  select(Name, Activity, Date, Time) %>%
  mutate(Time = str_sub(Time, 1,5))

cattle_stereotypic$Date <- format(cattle_stereotypic$Date, format = "%B %d, %Y")
  

formattable(cattle_stereotypic, align = c("l", rep("c",3)))

```


* Since there were only five observations in the data set where a cattle showed stereotypic behaviors, we decided to create a table instead of a visualization
* The table depicts the name of the cattle showing the stereotypic behavior, the type of stereotypic behavior, as well as the date and time of when the behavior occured 
* With only five cases, it appears as though cattles showing stereotypic behaviors is very rare 
* Also, of the three cattles at the zoo, only two of them were observed showing stereotypic behaviors


\

#### **Positive Behaviors**
*Is there one animal that spends more time doing the positive behaviors (Enrichment-based, Maintenance, Social, Locomotion)?*


```{r (CATTLE) Positive Behavior Visual}

#Creating Data Frame for Positive Behavior Visual
cattle_positive <- cattle_data %>% filter(Behavior_Type == "Positive") %>%
  group_by(Name, Category) %>% 
  summarize(Count = n()) 

cattle_positive <- cattle_positive %>% cbind(Percentages = cattle_positive$Count / rep(c(720, 761, 692), times = c(4,4,4)))

#Positive Behavior Visual
ggplot(data = cattle_positive) +
  geom_bar(aes(x = Name, y = Percentages, fill = Category), stat = "identity", width = .4) +
  labs(title = "Barplot of Positive Behaviors per Cattle",
       subtitle = "Percentages based on each cattle's total number of observations",
       x = "Cattle", y = "Percentage") +
  theme(plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 9, face = "italic"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L), 
                     limits = c(0,.4)) +
  scale_fill_manual(values = c("deeppink1", "springgreen3", "cyan2", "gold3")) 
  

```

* The visualization illustrates the percentage of times each cattle was observed doing a positive behavior
  + The percentages for each cattle are based on the total number of observations for that specific cattle
  + For example, considering **only** Batik's observations, roughly 36% is of those observations are of a positive behavior

* The visualization suggests that Batik spends the most time doing positive behaviors, followed by Patty, and then Chilly Jane
* The coloring indicates that the composition that makes up the total percentage of positive behaviors differs for each cattle
  + Chilly Jane spends a lot more time doing locomotion behaviors than maintanance behaviors
  + Patty spends a lot more time doing maintanence behaviors than locomotion behaviors
 
\


#### **Head Spin**
*Which animal is doing head spins and how often?*


```{r (MONKEY) Head spins Visual }

#Vector of total number of observations for each monkey

obs_per_monkey <- c(638,234,594,620,256,178)

#Barplot of Headspins (Percentage + Count)
ggplot(data = sq_monkey_data %>% filter(Activity == "Head spin")) +
  geom_bar(aes(x = Name, y= ..count.. / obs_per_monkey), fill = "turquoise3", width = .6) +
  labs(title = "Barplot of Head Spins per Monkey", 
       subtitle = "Percentages based on each monkey's total number of observations",
       x = "Monkey", y = "Percentage") +
  theme(plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 9, face = "italic")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L), 
                     limits = c(0,.125)) +
  geom_text(aes(x = Name, y = ..count.. / obs_per_monkey, label = ..count..), 
            stat = "count", vjust = -.4, size = 3.5, fontface = "italic")
```


* The visualization depicts the percentage of times each squirrel monkey was observed doing a head spin
  + The percentages for each monkey are based on the total number of observations for that specific monkey
  + For example, considering **only** Damian's observations, a little over 4% of those observations are of him doing a head spin. 
  
* The numbers on top of each bar represent the raw count of the number of head spins a particular animal was observed doing
  + Even though Filbert had 30 head spin observations while Damain only had 10, the percentage of head spins relative to their total number of observations is similar

* Notably, Brazil is observed doing a head spin on far more regular basis. Roughly 11% of all her observations are of her doing a head spin
* Isabel and Squirt are the monkeys that were observed doing a head spin least frequently relative to their total number of observations



\

#### **Goal Reflection**

1. Work with the Cattle data and answer the client's questions regarding the animal $\rightarrow$ **Complete**
    + We created visualizations and/or tables for each of the 4 questions regarding cattles that the client had
2. Work with Squirrel Monkey data and answer the client’s questions regarding the animal $\rightarrow$ **Complete**
    + We created visualizations and/or tables for each of the 3 questions regarding squirrel monkeys that the client had
3. Explore the differences in the structure of the animal datasets that we currently have $\rightarrow$ **Complete**
    + We created a table that highlighted the similarities and differences in structure between the three datasets
    + We also provided bullet point comments that provided further information
4. Create a prototype R Shiny program for only the Painted African Dogs, Cattles, and Squirrel Monkeys $\rightarrow$ **Complete**
    + We created a fully functioning prototype R Shiny program compatible with the three data sets currently avaliable to us
    + We published the program and posted a link so that people in STA-395 can test out the program
    + We also provided comments that describe the program in detail

\

* We were able to successfully complete all four of our goals
* All goals were feasible, but some goals took more time than others
* The first two goals were fairly straightforward, even though we were working with a completely new data set
* Since we had to explore the new data sets to complete the first two goals, the third goal was not troublesome
* The prototype R Shiny Program was the most complicated, and required a lot of preparation beforehand 


