---
title: "Pueblo Zoo Monitor Debriefing #3"
output: 
  html_document:
    theme: yeti
    toc: TRUE
    toc_depth: 5
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r Loading Libraries}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(lubridate)
library(gridExtra)
library(grid)
library(formattable)
library(zoo)
library(tibble)
library(forcats)
```


```{r Source}
source("dogs_cleaning.R")
source("cattle_cleaning.R")
source("sq_monkey_cleaning.R")
source("dogs_columns_addition.R")
source("cattle_columns_addition.R")
source("sq_monkey_columns_addition.R")
```

\

### **Goals**

1. **Work with the Cattle data and answer the client's questions regarding the animal**
    + Measured by the creation of visualizations to send to the client
2. **Work with another animal's data and answer the client’s questions regarding the animal**
    + Measured by the creation of visualizations to send to the client
3. **Explore the differences in the structure of the animal datasets that we currently have.**
    + Measured by the ability to articulate the differences during the next debriefing
4. **Create a prototype R Shiny program for only the Painted African Dogs, Cattles, and the other animal** 
    + Measured by a live demonstration during the next debriefing

\

### **Meeting with the Client**
We had our 3rd meeting with Anna on 3/18/2020 (Wednesday) to show some visuals and to interview her about the program.

#### **Plots** 
* We showed her 25 plots that answer her questions about Painted African Dogs, Cattles, and Squirrel Monkeys. (These plots include the ones we have already presented in our last debriefing)
* This part of the meeting was mainly about the clarification of the plots.
 
#### **Program** 
* We then asked her about her expectations for our final product, a program that will address all animals. Even though we had asked the same thing before, when we only had shown her some general plots, since we had created a wider variety of plots than then, we were curious if she had different ideas in mind. 

* Last time, she requested the three plots that we sent her at the first meeting (2/6) to be in the program:
  + Behaviors throught the day (Frequency & Proportion)
  + Activeness throught the week (Frequency)
  + Proportion of the behavior "Others" within all inactive behaviors

* We suggested that we create the input sections in the program so that the user can choose a factor by which they facet the data, e.g., day of week, time of day, animal's name, etc.
* She agreed on this suggestion and requested that we put some more detailed plots into the program, e.g., distribution of their observations, shift in animals' behaviors after a certain event, etc.

#### **Post Meeting**  

* Based on our client's request, we finalized the plots and choices of inputs we put in the prototype program:
  + Distribution of observations (by day of week/time of day/individual's name)
  + Percentage of a Behavior/Behaviors of users' choice per each individual animal
  + Percentage of a *Behavior Category/Behavior Categories of users' choice per each individual animal
  + Percentage of Behaviors before/after a certain event, e.g., death of the baby JT (in the Painted African Dogs' case)
  
*Behavior Category is a variable that groups some similar behaviors.
  
\

*Specific details regarding the program will be provided under Goal 4.* 

### **Goal 1**

***Visuals for Cattle***

\

#### **Agonistic Behaviors**
*How often are they showing Agonistic behaviors?*

\

#### **Stereotypic Behaviors**
*How often are they showing Stereoptypic behaviors?*

\

#### **Active vs. Inactive Behaviors**
*Are they spending more of their day active or inactive?*


```{r Visual for Cattle Active vs. Inactive for individual cattles}
##Common procedure
cattle_data$Hour <- as.numeric(cattle_data$Hour)
cattle_data$Category <- factor(cattle_data$Category, levels = c("Agonistic","Enrichment-based",
                                                                "Exploratory & Feeding", 
                                                                "Locomotion", "Maintenance", 
                                                                "Social", "Stereotypic", 
                                                                "Inactive", "Other"))
label_coloring <- rep(c("forestgreen","maroon"), times = c(7,2))

##plots ffor indivdual
cattle_data_grouped <- group_by(cattle_data, Name, Hour, Category)
summary <- as.data.frame(summarise(cattle_data_grouped, n()))
names(summary)[names(summary) == "n()"] <- "counts"
summary <- summary %>% group_by(Name, Hour) %>%
  mutate(sum = sum(counts))
cattle_data_HR <- left_join(cattle_data, summary, by = c("Name", "Hour", "Category"))
cattle_data_HR <- mutate(cattle_data_HR, percent = cattle_data_HR$counts/cattle_data_HR$sum*100)
cattle_data_HR$percent <- round(cattle_data_HR$percent, digits = 1)
cattle_data_HR$percent <- paste(cattle_data_HR$percent, "%")
##Plots of category by cattle by hour by name
#similar distribution
ggplot(cattle_data_HR, aes(x = Category, y = counts)) + 
  geom_bar(stat="identity", position=position_dodge(), fill = "salmon") +
  theme(axis.title.x = element_blank()) + 
  labs(title = "Bar Plot of Cattle's Behavior Category", subtitle = "Individuals", y="Counts") +
  facet_grid(Name ~ Hour) +
  theme(axis.text.x = element_text(angle = 90, color = label_coloring))
```

```{r Visual for Cattle Active vs. Inactive for all cattles}
cattle_data_grouped2 <- group_by(cattle_data, Hour, Category)
summary2 <- as.data.frame(summarise(cattle_data_grouped2, n()))
names(summary2)[names(summary2) == "n()"] <- "counts"
summary2 <- summary2 %>% group_by(Hour) %>%
  mutate(sum = sum(counts))
cattle_data_HR2 <- left_join(cattle_data, summary2, by = c("Hour", "Category"))
cattle_data_HR2 <- mutate(cattle_data_HR2, percent = cattle_data_HR2$counts/cattle_data_HR2$sum*100)
cattle_data_HR2$percent <- round(cattle_data_HR2$percent, digits = 1)
cattle_data_HR2$percent <- paste(cattle_data_HR2$percent, "%")
ggplot(cattle_data_HR2, aes(x = Category, y = counts)) + 
  geom_bar(stat="identity", position=position_dodge(), fill = "salmon") +
  labs(x = "Behavior Category", subtitle = "General", y="Counts") +
  facet_grid(. ~ Hour) +
  theme(axis.text.x = element_text(angle = 90, color = label_coloring))
```

* Green = Active, Red = Inactive

\

* Cattles generally spend more actively around 10am and 11am and more inactively after lunch time
* All individuals behave similarly throuout the day
* Significantly few observations at 12pm and 4pm, as in the Painted African Dogs' case

\

#### **Positive Behaviors**
*Is there one animal that spends more time doing the positive behaviors (Enrichment-based, Maintenance, Social, Locomotion)?*

\

### **Goal 2**

***Visuals for Squirrel Monkeys***

\

#### **Aggressive Behaviors**
*Is it one squirrel monkey being aggressive to all the others or is it one being picked on regularly?*


\

#### **Headspin**
*Which animal is doing headspins and how often?*


\

#### **Impact of Events on Behaviors**
*Did the deaths of Damian and Pistachio have an impact on the monkeys' behaviors?*

*Did the addition of Squirt have an impact on the monkeys' behaviors?*

```{r Visual for before and after the death of Damian}
##Damian's death
Damian_before <- subset(sq_monkey_data, Date < "2018-02-23")
Damian_before <- Damian_before %>% group_by(Activity)
summary_Damian_before <- as.data.frame(summarise(Damian_before, n()))
names(summary_Damian_before)[names(summary_Damian_before) == "n()"] <- "counts"
summary_Damian_before <- summary_Damian_before %>%
  mutate(percent = round(counts/399*100, 1)) %>%
  mutate(Period = "Before")

Damian_amonth <- subset(sq_monkey_data, Date >= "2018-02-23" & Date <= "2018-03-21")
Damian_amonth <- Damian_amonth %>% group_by(Activity)
summary_Damian_amonth <- as.data.frame(summarise(Damian_amonth, n()))
names(summary_Damian_amonth)[names(summary_Damian_amonth) == "n()"] <- "counts"
summary_Damian_amonth <- summary_Damian_amonth %>%
  mutate(percent = round(counts/159*100, 1)) %>%
  mutate(Period = "One month after")

Damian_after <- subset(sq_monkey_data, Date > "2018-03-21")
Damian_after <- Damian_after %>% group_by(Activity)
summary_Damian_after <- as.data.frame(summarise(Damian_after, n()))
names(summary_Damian_after)[names(summary_Damian_after) == "n()"] <- "counts"
summary_Damian_after <- summary_Damian_after %>%
  mutate(percent = round(counts/1962*100, 1)) %>%
  mutate(Period = "After")

summary_Damian <- rbind(summary_Damian_before, summary_Damian_amonth, summary_Damian_after)
summary_Damian$Period <- factor(summary_Damian$Period, levels = c("Before", "One month after", "After"))
  
ggplot(summary_Damian, aes(x="", y=percent, fill=fct_reorder(Activity, desc(percent)))) + geom_bar(stat="identity", width=1) +
  facet_grid(.~ Period) +
  coord_polar("y", start=0) + 
  labs(x = NULL, y = NULL, fill = NULL, title = "Damian's Passing and Activity", subtitle = "Raw Counts: Before = 399, One month after = 159, After = 1962") +
  guides(fill = guide_legend(reverse = TRUE, override.aes = list(size = 1))) +
  theme_classic() + theme(axis.line = element_blank(),
                          axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          plot.title = element_text(hjust = 0.5, face = "bold"),
                          plot.subtitle = element_text(face = "italic"),
                          legend.position="bottom") +
  scale_fill_manual(values = rainbow(16)[sample(1:16)])
```

```{r Visual for before and after the death of Pistachio}
##Pistachio's death
Pistachio_before <- subset(sq_monkey_data, Date < "2018-12-03")
Pistachio_before <- Pistachio_before %>% group_by(Activity)
summary_Pistachio_before <- as.data.frame(summarise(Pistachio_before, n()))
names(summary_Pistachio_before)[names(summary_Pistachio_before) == "n()"] <- "counts"
summary_Pistachio_before <- summary_Pistachio_before %>%
  mutate(percent = round(counts/760*100, 1)) %>%
  mutate(Period = "Before")

Pistachio_after <- subset(sq_monkey_data, Date >= "2018-12-03")
Pistachio_after <- Pistachio_after %>% group_by(Activity)
summary_Pistachio_after <- as.data.frame(summarise(Pistachio_after, n()))
names(summary_Pistachio_after)[names(summary_Pistachio_after) == "n()"] <- "counts"
summary_Pistachio_after <- summary_Pistachio_after %>%
  mutate(percent = round(counts/1760*100, 1)) %>%
  mutate(Period = "After")

summary_Pistachio <- rbind(summary_Pistachio_before, summary_Pistachio_after)
summary_Pistachio$Period <- factor(summary_Pistachio$Period, levels = c("Before", "After"))

ggplot(summary_Pistachio, aes(x="", y=percent, fill=fct_reorder(Activity, desc(percent)))) + geom_bar(stat="identity", width=1) +
  facet_grid(.~ Period) +
  coord_polar("y", start=0) + 
  labs(x = NULL, y = NULL, fill = NULL, title = "Pistachio's Passing and Activity", subtitle = "Raw Counts: Before = 760, After = 1760") +
  guides(fill = guide_legend(reverse = TRUE, override.aes = list(size = 1))) +
  theme_classic() + theme(axis.line = element_blank(),
                          axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          plot.title = element_text(hjust = 0.5, face = "bold"),
                          plot.subtitle = element_text(face = "italic"),
                          legend.position="bottom") +
  scale_fill_manual(values = rainbow(27)[sample(1:27)])
```

```{r Visual for before and after the joining of Squirt}
##Squirt's joining
Squirt_before <- subset(sq_monkey_data, Date < "2019-06-13")
Squirt_before <- Squirt_before %>% group_by(Activity)
summary_Squirt_before <- as.data.frame(summarise(Squirt_before, n()))
names(summary_Squirt_before)[names(summary_Squirt_before) == "n()"] <- "counts"
summary_Squirt_before <- summary_Squirt_before %>%
  mutate(percent = round(counts/1296*100, 1)) %>%
  mutate(Period = "Before")

Squirt_after <- subset(sq_monkey_data, Date >= "2019-06-13")
Squirt_after <- Squirt_after %>% group_by(Activity)
summary_Squirt_after <- as.data.frame(summarise(Squirt_after, n()))
names(summary_Squirt_after)[names(summary_Squirt_after) == "n()"] <- "counts"
summary_Squirt_after <- summary_Squirt_after %>%
  mutate(percent = round(counts/1224*100, 1)) %>%
  mutate(Period = "After")

summary_Squirt <- rbind(summary_Squirt_before, summary_Squirt_after)
summary_Squirt$Period <- factor(summary_Squirt$Period, levels = c("Before", "After"))

ggplot(summary_Squirt, aes(x="", y=percent, fill=fct_reorder(Activity, desc(percent)))) + geom_bar(stat="identity", width=1) +
  facet_grid(.~ Period) +
  coord_polar("y", start=0) + 
  labs(x = NULL, y = NULL, fill = NULL, title = "Squirt's Joining and Activity", subtitle = "Raw Counts: Before = 1296, After = 1224") +
  guides(fill = guide_legend(reverse = TRUE, override.aes = list(size = 1))) +
  theme_classic() + theme(axis.line = element_blank(),
                          axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          plot.title = element_text(hjust = 0.5, face = "bold"),
                          plot.subtitle = element_text(face = "italic"),
                          legend.position="bottom") +
  scale_fill_manual(values = rainbow(27)[sample(1:27)])
```

* Deaths/Joining information
  + Damian passed away on Feb 22, 2018,
  + Pistachio passed away on Dec 3, 2018
  + Squirt came to the zoo on June 13, 2019
* Included the raw counts of observations since they normally differ significantly between before/after
* Only Damian has the period "One month after" since only he had observations soon after his death.
* We also created the same plots for Painted African Dogs' death/birth and sent them to our client since it was one of her initial interests. 

\

* We see a larger proportion of "Sitting" after a certain event, but not clear what this explains.

\

### **Goal 3**

***Comparison among animal datasets***


### **Goal 4**

***R Shiny app Prototype***

##link to it

\

### **Reflection and Discussion**

\

#### **Progress Reflection**

**1. Things that went well**

* Effective work distribution (using Trello)
* Effective use of R documents; we created an individual, but shared, R scripts to avoid the possible merge conflicts on Github
* Creation of the visuals
  + 25 visuals (Including some created for Debriefing #2) were sent to our client
  + She liked the plots and commented on some of them; expecially, she mentioned that the weather/temperature data would not be too useful since those two factors might be different from those at the recorded place (Pueblo Airport)
* Creation of the generalized cleaning script
  + Created a script that is compatible with any animal's dataset
  + This script takes the differences among the datasets into account and cleans any dataset in the same way 
* Meeting with our client 
  + She told us her more specific preferences on the plots that would possibly be included in the program than the last meeting
  + We were able to envision the program soon after the meeting and decide the plots and inputs for each plot 

**2. Things that were difficult**

* Plot creations
  + It took time to figure out how to deal with the cases with extremely few observations
  + Ended up creating a table for those ones
* Pie chart aesthetics
  + We had to choose between making the colors of the slice distinct to each other and sorting the slices by size (proportion)
  + Ended up making the colors of the slice distinct to each other and putting the largest slice at the beginning
  
\

#### **Goal Reflection**

\

#### **Next Steps**









