---
title: "Pueblo Zoo Monitor Debriefing #2"
date: "3/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r Loading Libraries}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(lubridate)
library(gridExtra)
library(grid)
library(formattable) #To make a nice table
library(zoo)
library(tibble)
library(ggbeeswarm)
```


```{r Source}
source("cleaning.R")
source("columns_addition.R")
```

\

#### **Goals**

\

\

\

#### **Meeting with the Client**


\

* **Questions**  
  + Level of technology and whether they want to use R (data cleaning procedure report)
  + Format of the instruction document (Google Docs or R markdown)  
  + Expectation for the program (Visuals etc. )  
  + How many/which animals in the program?  
  + Data structure consistency  
  + Data storage    

\

* **Answers**  
  + They're fairly unfamiliar with the data analysis tools/technologies beyond Excel, but willing to adopt R as a way to clean data in the future  
  + R markdown with both codes and an instruction  
  + Three plots that we created:  
    $\rightarrow$ Active vs. Inactive proportion throughout the week  
    $\rightarrow$ Behavior frequencies throughout the day  
    $\rightarrow$ Proportion of the inactive behaviors  
  + Ideally all animals, 45 species in total 
  + The data structure is fundamentally the same but there could be more columns depending on the animal's activeness
  + ZooMonitor (not too troublesome to transfer the data to Google Drive)

\



#### **Data**

\

\

\

#### **Goal 1**

\

\

\

#### **Goal 2**

\

\

\


#### **Goal 3**

\

\

\


#### **Goal 4**


\

\

\


#### **Progress Reflection**

\

\

\

#### **Goal Reflection**

\

\

\


#### **Next Steps**

\

\

\











