---
title: 'Progress Report #2'
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Loading Libraries}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(lubridate)
library(gridExtra)
library(grid)
library(formattable) #To make a nice table
library(zoo)
library(tibble)
library(ggbeeswarm)
```


```{r Source}
source("cleaning.R")
source("columns_addition.R")
```


```{r Inactivity Plot in 6 Month Intervals}

```



```{r Percentage of Observations based on Time of Day/Day of Week}
time_of_day_viz <- ggplot(data = dogs_data, aes(x = Hour)) + 
  geom_bar(aes(y = ..count../nrow(dogs_data)*100), fill = "steelblue", width = .75) + 
  scale_x_discrete(limits = 9:16) +
  scale_y_continuous(limits = c(0,25)) + 
  geom_hline(yintercept = (1/8)*100, color = "darkmagenta", alpha = .45, linetype = "longdash") +
  labs(title = "Percentage of Observations (Per Time of Day)", x = "Time of Day", y = "Percentage (%)") +
  annotate("text", x= 16.6, y = 13.5 , label = "12.5%", color = "darkmagenta", size = 3.25)


day_of_week_viz <- ggplot(data = dogs_data, aes(x = Day_of_Week)) +
  geom_bar(aes(y = ..count../nrow(dogs_data)*100), fill = "steelblue2", width = .75) +
  scale_x_discrete(limits=c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")) +
  scale_y_continuous(limits = c(0,60)) +
  labs(title = "Percentage of Observations (Per Day of Week)", x = "Day of Week", y = "Percentage (%)") +
  geom_hline(yintercept = (1/7)*100, color = "darkmagenta", alpha = .45, linetype = "longdash") + 
  annotate("text", x= 1.5, y = 16.5 , label = "14.28%", color = "darkmagenta", size = 3.25)

grid.arrange(day_of_week_viz, time_of_day_viz,  nrow = 1)

```

+ The plot on the left illustrates the percentage of observations recorded on each day of the week. 
+ If observations were recorded equally throughout the days of the week, then the height of the bars should match the purple dashed line.
+ The plot suggests that over 50% of all observations in the data are recorded on Saturdays.
+ Friday is the day of the week where the least number of observations were recorded. 


+ The plot on the right displays the percentage of observations recorded on each hour of the day. + If equal numbers of observations were recorded on each hour of the day, the height of the bars should match up with the purple dashed line. 
+ The plot suggests that 9AM, 10AM, and 2PM (14), are the times in which most observations are taken.
+ It appears as though not a lot of observations are recorded at 12PM and 4PM (16).









