---
title: 'Progress Report #2'
author: "Shuyi Qi, Hinako Minagi, Riu Sakaguchi"
date: "2/27/2020"
output: 
  html_document:
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r Loading Libraries}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(lubridate)
library(gridExtra)
library(grid)
library(formattable) #To make a nice table
library(zoo)
library(tibble)
library(ggbeeswarm)
```


```{r Source}
source("cleaning.R")
source("columns_addition.R")
```

#### **Notes**

Hello Anna! We have created some visualizations that you requested during the first meeting and in the follow-up email.  
The figures include:  

+ **Figure 1**: Barplots displaying the frequency and percentage of dogs' active and inactive behaviors based on the six-month intervals. This plot helps you discover the shifts in the dogs' behaviors throughout the observational period.  
+ **Figure 2**: Barplots displaying the percentage of observations per day of week, and time of day. This plot is useful in identifying certian days of the week or times of the day where few/many observations have been recorded.

We hope that the visualizations we created in this progress report is helpful for you! 


<br>

#### **Figure 1.**

```{r Inactivity Plot in 6 Month Intervals, fig.height = 15, fig.width = 12}
## Frequency and percentage of active and inactive behaviors (in six-month intervals)
## Calculate percentage for each behavior within each six-month interval
active_obs <- dogs_data %>% 
  filter(Activeness == 'Active') %>% 
  group_by(Six_Month_Interval, Activity) %>%
  summarize(n = n()) %>%
  mutate(Percentage = n/sum(n)*100)

## Calculate percentage for each inactive behavior within each six-month interval
inactive_obs <- dogs_data %>% 
  filter(Activeness == 'Inactive') %>% 
  group_by(Six_Month_Interval, Activity) %>%
  summarize(n = n()) %>%
  mutate(Percentage = n/sum(n)*100)

active_percentage <- ggplot(active_obs, aes(x=Six_Month_Interval, y=Percentage, fill=Activity)) +
  geom_col(position = "stack", width = 0.4) +
  labs(title = "b. Percentage of active behaviors (in 6-month intervals)", x = "Six Month Intervals", y = "Percentage of observations (%)") + 
  scale_fill_manual(values = c("coral","cornflowerblue", "darkolivegreen3", "darkorchid1", "darkcyan")) + 
  theme(plot.title = element_text(face = "bold"))

inactive_percentage <- ggplot(inactive_obs, aes(x=Six_Month_Interval, y=Percentage, fill=Activity)) +
  geom_col(position = "stack", width = 0.4) +
  labs(title = "d. Percentage of inactive behaviors (in 6-month intervals)", x = "Six Month Intervals", y = "Percentage of observations (%)") +
  scale_fill_manual(values = c("coral","cornflowerblue", "darkolivegreen3", "darkorchid1", "darkcyan")) +
  theme(plot.title = element_text(face = "bold"))

## Create barplots for the frequencies
active_obs2 <- dogs_data %>% filter(Activeness == 'Active')
inactive_obs2 <- dogs_data %>% filter(Activeness == "Inactive")
active_frequency <- ggplot(data=active_obs2) + 
  geom_bar(aes(x=Activity), fill = "coral", alpha = 0.7) + 
  labs(title = "a. Frequency of active behaviors (in 6-month intervals)", x = "Dogs' behavior when active", y = "Frequency") +
  facet_grid(. ~Six_Month_Interval) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), plot.title = element_text(face = "bold"))

inactive_frequency <- ggplot(data=inactive_obs2) + 
  geom_bar(aes(x=Activity), fill = "coral", alpha = 0.7) + 
  labs(title = "c. Frequency of inactive behaviors (in 6-month intervals)", x = "Dogs' behavior when inactive", y = "Frequency") +
  facet_grid((. ~Six_Month_Interval)) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), plot.title = element_text(face = "bold"))

grid.arrange(active_frequency, active_percentage, inactive_frequency, inactive_percentage, nrow = 4)

```

* Plot **1a** and plot **1b** illustrate the frequency and percentage of different types of **active** behaviors recorded in six-month intervals.
* 'Running' is the most common active behavior, especially between 2018-09-01 and 2019-03-01. 'Walking' is the second most common active behavior, especially between 2019-09-01 and 2020-01-11.

* Plot **1c** and plot **1d** illustrate the frequency and percentage of different types of **inactive** behaviors recorded in six-month intervals.
* The total number of observations for inactivity between 2019-03-01 and 2020-01-11 is significantly higher and the majority of these observations are 'Resting'.

* We notice that the total number of observations is not evenly distributed across the four six-month intervals. Specifically, there are more observations for both activity and inactivity between 2019-03-01 and 2019-09-01. 

\

\

\

<br>

#### **Figure 2.**

```{r Percentage of Observations based on Day of Week/Time of Day, fig.height = 6, fig.width = 10}
time_of_day_viz <- ggplot(data = dogs_data, aes(x = Hour)) + 
  geom_bar(aes(y = ..count../nrow(dogs_data)*100), fill = "steelblue", width = .75) + 
  scale_x_discrete(limits = 9:16) +
  scale_y_continuous(limits = c(0,25)) + 
  geom_hline(yintercept = (1/8)*100, color = "darkmagenta", alpha = .45, linetype = "longdash") +
  labs(title = "b. Percentage of Observations (Per Time of Day)", x = "Time of Day", y = "Percentage (%)") +
  annotate("text", x= 16.6, y = 13.5 , label = "12.5%", color = "darkmagenta", size = 3.25) +
  theme(plot.title = element_text(face = "bold"))

day_of_week_viz <- ggplot(data = dogs_data, aes(x = Day_of_Week)) +
  geom_bar(aes(y = ..count../nrow(dogs_data)*100), fill = "steelblue2", width = .75) +
  scale_x_discrete(limits=c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")) +
  scale_y_continuous(limits = c(0,60)) +
  labs(title = "a. Percentage of Observations (Per Day of Week)", x = "Day of Week", y = "Percentage (%)") +
  geom_hline(yintercept = (1/7)*100, color = "darkmagenta", alpha = .45, linetype = "longdash") +   annotate("text", x= 1.5, y = 16.5 , label = "14.28%", color = "darkmagenta", size = 3.25) +
  theme(plot.title = element_text(face = "bold"))

grid.arrange(day_of_week_viz, time_of_day_viz,  nrow = 1)

```

**2a.**

* The plot on the **left** illustrates the percentage of observations recorded on each day of the week. 
* If observations were recorded equally throughout the days of the week, then the height of the bars should match the purple dashed line.
  + That is, each day would consist of 14.28% of the total number of observations if there was equal representation for the days of the week.
* The plot suggests that over 50% of all observations in the data are recorded on Saturdays.
* Friday is the day of the week where the least number of observations were recorded. 


**2b.**

* The plot on the **right** displays the percentage of observations recorded on each hour of the day.
* If equal numbers of observations were recorded on each hour of the day, the height of the bars should match up with the purple dashed line. 
  + In otherwords, each hour of the day would consist of 12.5% of the total number of observations if there was equal representation for the times of the day.
* The plot suggests that 9AM, 10AM, and 2PM (14), are the times in which most observations are recorded.
* It appears as though not a lot of observations are recorded at 12PM and 4PM (16).

\


